---
title: "Supplementary Materials"
shorttitle: Supplementary Materials
output: 
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: true
    keep_tex: true
#classoption: man
header-includes:
- \usepackage{float}
- \usepackage{placeins}
- \usepackage{setspace}
- \usepackage{subcaption}
- \usepackage{array}
- \makeatletter
- \makeatother
- \raggedbottom
- \setlength{\parindent}{0.5in}
---

```{r include= FALSE}
knitr::opts_chunk$set(warning = FALSE, cache = FALSE)
knitr::opts_chunk$set(engine.opts = list(verbatim = TRUE))
```

\clearpage

# Code for the Illustrative Example

```{r message = FALSE}
# Load libraries
library(here)
library(dplyr)
library(kableExtra)
library(lavaan)
# install.packages("remotes")
# remotes::install_github("marklhc/pinsearch")
library(pinsearch)
#remotes::install_github("mmm-usc/unbiasr", ref = "pkg")
library(unbiasr)
source('code/helper_functions_for_tables.R')
```

```{r include = FALSE}
# This chunk determines whether to rerun some of the chunks to recreate the analyses
# reported in the main manuscript, including the tables and figures.
if (!dir.create("rds")) dir.create("rds")
eval_dat <- !file.exists(here("addhealth_w1_dat.rds"))
eval_config_fit <- !file.exists(here("rds/configural_fit_4g.rds"))

# dat <- readRDS("rds/addhealth_w1_dat_updated.rds")

# Configural and partial fits
# configural_fit_W <- readRDS("rds/configural_fit_2g_W.rds")
# configural_fit_H <- readRDS("rds/configural_fit_2g_H.rds")

eval_pin_out <- !file.exists("rds/pin_out.rds")
eval_pin_out_W <- !file.exists("rds/pin_out_W.rds")
eval_pin_out_H <- !file.exists("rds/pin_out_H.rds")
```

## Data preparation

```{r eval = eval_dat, message=FALSE}
# Add Health Wave I, 1994-1995 public use data 
# https://dataverse.unc.edu/dataset.xhtml?persistentId=doi:10.15139/S3/11900
# Put the downloaded files in a folder named 'dataverse_files' in the working directory

dt <- haven::read_sas(here("dataverse_files/w1inhome.sas7bdat")) 
# See 'pubwgt1.pdf' for more info on the public use weights
weights <- haven::read_sas(here("dataverse_files/w1weight.sas7bdat"))

CESD_vars <- c(do.call(paste0, expand.grid("H1FS", 1:19)))
cesd_labs <- 
  c("bothered", "appetite", "blues", "good", "mind", "depressed", "tired", "hopeful",
    "failure", "fearful", "happy", "talk", "lonely", "unfriendly", "enjoyed", "sad",
    "dislike", "getstarted", "life")
CESD_n <- c(do.call(paste0, expand.grid("cesd", c(1:19))))

# create an ethnicity variable with levels "Black", 
# "White", "Hispanic", and "Asian" using Boolean operators 
dt$ethnicity <- "Other"
dt[dt$H1GI6B==1 & dt$H1GI6A==0 & dt$H1GI6D==0 & dt$H1GI4==0, "ethnicity"] <-"Black"
dt[dt$H1GI6B==0 & dt$H1GI6A==1 & dt$H1GI6D==0 & dt$H1GI4==0, "ethnicity"] <-"White"
dt[dt$H1GI6B==0 & dt$H1GI6A==0 & dt$H1GI6D==1 & dt$H1GI4==0, "ethnicity"] <-"Asian"
dt[dt$H1GI6B==0 & dt$H1GI6A==0 & dt$H1GI6D==0 & dt$H1GI4==1, "ethnicity"] <-"Hispanic"

dt <- dt %>% select(AID, BIO_SEX, ethnicity, all_of(CESD_vars))
colnames(dt) <- c("AID", "sex", "ethnicity", CESD_n)

# Select the subset with complete CESD scores (filter rows where no item was 
# 'refused' or 'don't know') and filter out the Other groups
dt_sub <- dt %>%  
  filter_at(vars(starts_with('cesd')), all_vars(. != 6 & . != 8)) %>% # removes 47 obs. 
  filter(ethnicity != "Other") 
colnames(dt_sub) <- c("AID", "sex", "ethnicity", cesd_labs) # 5848

# Merge with weights
dat <- left_join(dt_sub, weights, "AID") # 5848 x 24

# Recode positively worded items to achieve a 'lack of' interpretation for consistency 
dat$happy <- recode(dat$happy, `0` = 3, `1` = 2, `2` = 1, `3` = 0)
dat$good <- recode(dat$good, `0` = 3, `1` = 2, `2` = 1, `3` = 0)
dat$hopeful <- recode(dat$hopeful, `0` = 3, `1` = 2, `2` = 1, `3` = 0)
dat$enjoyed <- recode(dat$enjoyed, `0` = 3, `1` = 2, `2` = 1, `3` = 0)

labs <- paste0("i", 1:21)
labs[7] <- "i7r"; labs[20] <- "i20r"
labs <- labs[-c(11, 17)]
colnames(dat) <- c("AID", "sex", "ethnicity", labs, "CLUSTER2", "GSWGT1")

# Define binary ethnicity variables for the examples where groups are collapsed
dat$White <- "non-White"
dat$White[dat$ethnicity=="White"] <- "White"
dat$Hispanic <- "non-Hispanic"
dat$Hispanic[dat$ethnicity=="Hispanic"] <- "Hispanic"
```

```{r eval = eval_dat, include = FALSE}
saveRDS(dat, here("addhealth_w1_dat.rds"))
```

```{r}
dat <- readRDS("addhealth_w1_dat.rds")
```

```{r include = FALSE, eval = TRUE}
labs <- paste0("i", 1:21)
labs[7] <- "i7r"; labs[20] <- "i20r"
labs <- labs[-c(11, 17)]
```

```{r message=FALSE, eval = TRUE}
# Define miscellaneous variables
item_factor <- paste0(
  labs, "-",  c("F1", "F1", "F2", "F3", "F1", "F2", "F1", "F3", "F2", "F2", "F3", "F1", 
       "F2", "F4", "F3", "F2", "F4", "F1", "F2" ))
groups_wbha <- c("White", "Black", "Hispanic", "Asian")
fitv <- c("chisq", "npar", "df", "pvalue", "rmsea.robust", "cfi.robust", "tli.robust", 
 "srmr")
cols <- c("#FF5733", "springgreen2", "magenta", "#33C3Ff")
colsW <- c(cols[1], "black")
colsH <- c("black", cols[3])
names(colsW) <- c("White", "non-White")
```


\clearpage
## Multi-group MCAA (4 groups)

```{r}
# Define the configural model string
m1 <- '
 F1 =~ i1 + i2 + i5 + i7r + i13 + i20r
 F2 =~ i3 + i6 + i9 + i10 + i14 + i18 + i21
 F3 =~ i4 + i8 + i12 + i16
 F4 =~ i15 + i19
'
```

```{r eval = eval_config_fit, message=FALSE}
# Fit the configural model
configural_fit <- cfa(
  model = m1, data = dat, group = "ethnicity", estimator = "MLR", missing = "fiml", 
  se = "robust.mlr", sampling.weights = "GSWGT1", group.label = groups_wbha) 
```

```{r eval = eval_config_fit}
saveRDS(configural_fit, "rds/configural_fit_4g.rds")
```

```{r include = FALSE}
configural_fit <- readRDS("rds/configural_fit_4g.rds")
```

```{r eval = eval_pin_out, message=FALSE}
# Use pinsearch to automate the partial invariance model search process
pin_out <- pinSearch(
  config_mod = m1, data = dat, group = "ethnicity", estimator = "MLR", missing = "fiml", 
  se = "robust.mlr", sampling.weights = "GSWGT1", group.label = groups_wbha, 
  type = "residuals")
partial_fit <- pin_out[[1]]
```

```{r eval = eval_pin_out}
saveRDS(pin_out, "rds/pin_out.rds")
```

```{r include = FALSE}
pin_out <- readRDS("rds/pin_out.rds")
partial_fit <- pin_out[[1]]
```

```{r message=FALSE}
# Equivalently, arguments used in repeated function calls can be stored in lists and 
# passed to do.call() as below to minimize code repetition
args_cfa <- 
  list(data = dat, group = "ethnicity", estimator = "MLR", missing = "fiml", 
       se = "robust.mlr", sampling.weights = "GSWGT1", group.label = groups_wbha)
```

```{r eval = FALSE}
configural_fit <- do.call(cfa, c(m1, args_cfa))
```

```{r}
args_pin <- c(args_cfa, type = "residuals")
```

```{r eval = FALSE}
pin_out <- do.call(pinSearch, c(m1, args_pin))
partial_fit <- pin_out[[1]]
```

```{r message=FALSE}
# Extract model fit indices
config_fit_stat <- fitMeasures(configural_fit, fitv)
partial_fit_stat <- fitMeasures(partial_fit, fitv)

# Examine fMACS effect size indices reflecting item-level bias across groups
es <- pinsearch::pin_effsize(partial_fit)
sort(round(es[,item_factor[item_factor %in% colnames(es)]], 3), decreasing = TRUE)

# Compute mixing proportions
weightsums <- tapply(dat$GSWGT1, dat$ethnicity, sum)
mixp_unord <- weightsums / sum(weightsums)
mixp <- matrix(c(mixp_unord[4], mixp_unord[2], mixp_unord[3], mixp_unord[1]), ncol = 4)
colnames(mixp) <- groups_wbha
```

```{r message=FALSE, fig.show='hide'}
# Latent factor weights: place weights proportional to item counts
lw <- c(6, 7, 4, 2) / sum(c(6, 7, 4, 2))

# Call PartInv() using the cfa() output object
out_4g <- PartInv(
  cfa_fit = partial_fit, pmix = mixp, weights_latent = lw, cut_z = 15.2, 
  plot_contour = TRUE, show_mi_result = TRUE, custom_colors = cols)
```

```{r include = TRUE}
# Overall ES
to_asin <- function(x) 2 * asin(sqrt(x))
get_multi_h <- function(pvec) {
  asin_pvec <- to_asin(pvec)
  2 * sqrt(mean((asin_pvec - mean(asin_pvec))^2))
}
multi_h_4g <- apply(out_4g$summary[, c(1, 5:7)], MARGIN = 1, FUN = get_multi_h)
multi_h_4g
```

```{r fig.show='hide'}
# Alternatively, call PartInv() using lists of parameter estimates 
est <- unnest_list(lavInspect(partial_fit, "est"))
out_4g <- PartInv(
  alpha = est$alpha, psi = est$psi, lambda = est$lambda, nu = est$nu, 
  theta = est$theta, pmix = mixp, weights_latent = lw, cut_z = 15.2, 
  plot_contour = TRUE, show_mi_result = TRUE, custom_colors = cols,
  labels = groups_wbha)
```

```{r fig.show='hide'}
# Plot CAI and Adverse Impact Ratio across a range of thresholds
plot_CAI_across_range(
  x = out_4g, custom_colors = cols, cutoffs_from = 10, cutoffs_to = 30, 
  add_vertical_threshold_at = 15.2)
```

\clearpage
## Multi-group MCAA (two groups)
### White vs. non-White
```{r message=FALSE}
# Update stored values, call do.call() to obtain configural and partial invariance models
args_cfa_W <- args_cfa
args_pin_W <- args_pin
args_pin_W$group <- args_cfa_W$group <- "White"
args_pin_W$group.label <- args_cfa_W$group.label <- c("non-White", "White")

configural_fit_W <- do.call(cfa, c(m1, args_cfa_W))
```

```{r eval = eval_pin_out_W}
pin_out_W <- do.call(pinSearch, c(m1, args_pin_W))
partial_fit_W <- pin_out_W[[1]]
```

```{r eval = eval_pin_out_W}
saveRDS(pin_out_W, "rds/pin_out_W.rds")
```

```{r include = FALSE}
pin_out_W <- readRDS("rds/pin_out_W.rds")
partial_fit_W <- pin_out_W[[1]]
```

```{r}
# Examine fMACS effect size indices reflecting item-level bias across groups
es_W <- pinsearch::pin_effsize(partial_fit_W)
sort(round(es_W[, item_factor[item_factor %in% colnames(es_W)]], 3), decreasing = TRUE)

config_fit_stat_W <- fitMeasures(configural_fit_W, fitv)
partial_fit_stat_W <- fitMeasures(partial_fit_W, fitv)

weightsums_W <- tapply(dat$GSWGT1, dat$White, sum)
mixp_W <- weightsums_W / sum(weightsums_W)
```

```{r message=FALSE, echo=TRUE, fig.show='hide'}
out_2g_W <- PartInv(
  cfa_fit = partial_fit_W, pmix = mixp_W, weights_latent = lw,
  cut_z = 15.2, plot_contour = TRUE, show_mi_result = TRUE, 
  custom_colors = colsW, reference = "White")

plot_CAI_across_range(
  x = out_2g_W, cutoffs_from = 10, cutoffs_to = 30, 
  add_vertical_threshold_at = 15.2, custom_colors = colsW)
```



### Hispanic vs. non-Hispanic
```{r message=FALSE}
# Update stored values, call do.call() to obtain configural and partial invariance models
args_cfa_H <- args_cfa
args_pin_H <- args_pin
args_pin_H$group <- args_cfa_H$group <- "Hispanic"
args_pin_H$group.label <- args_cfa_H$group.label <- c("non-Hispanic", "Hispanic")

configural_fit_H <- do.call(cfa, c(m1, args_cfa_H))
```

```{r eval = eval_pin_out_H}
pin_out_H <- do.call(pinSearch, c(m1, args_pin_H))
partial_fit_H <- pin_out_H[[1]]
```

```{r eval = eval_pin_out_H}
saveRDS(pin_out_H, "rds/pin_out_H.rds")
```

```{r include = FALSE}
pin_out_H <- readRDS("rds/pin_out_H.rds")
partial_fit_H <- pin_out_H[[1]]
```

```{r}
# Examine fMACS effect size indices reflecting item-level bias across groups
es_H <- pinsearch::pin_effsize(pin_out_H[[1]])
sort(round(es_H[, item_factor[item_factor %in% colnames(es_H)]], 3), decreasing = TRUE)

config_fit_stat_H <- fitMeasures(configural_fit_H, fitv)
partial_fit_stat_H <- fitMeasures(partial_fit_H, fitv)

weightsums_H <- tapply(dat$GSWGT1, dat$Hispanic, sum)
mixp_unord <- weightsums_H / sum(weightsums_H)
mixp_H <- matrix(c(mixp_unord[2], mixp_unord[1]), ncol = 2)
colnames(mixp_H) <- c("non-Hispanic", "Hispanic")
```


```{r message=FALSE, echo=TRUE, fig.show='hide'}
out_2g_H <- PartInv(
  cfa_fit = partial_fit_H, pmix = mixp_H, weights_latent = lw, cut_z = 15.2, 
  plot_contour = TRUE, show_mi_result = TRUE, custom_colors = colsH, 
  reference = "non-Hispanic")

plot_CAI_across_range(
  x = out_2g_H, cutoffs_from = 10, cutoffs_to = 30, by = 0.1,
  add_vertical_threshold_at = 15.2, custom_colors = colsH)
```



```{r include = FALSE}
### RUN TO SAVE RDS FILES ###
saveRDS(configural_fit_W, "rds/configural_fit_2g_W.rds")
saveRDS(configural_fit_H, "rds/configural_fit_2g_H.rds")
saveRDS(mixp, "rds/mixp_4g_wbha.rds")
saveRDS(mixp_W, "rds/mixp_2g_W.rds")
saveRDS(mixp_H, "rds/mixp_2g_H.rds")
saveRDS(lw, "rds/lw.rds")

saveRDS(out_4g, "rds/PartInv_out_4g.rds")
saveRDS(out_2g_W, "rds/PartInv_out_2g_W.rds")
saveRDS(out_2g_H, "rds/PartInv_out_2g_H.rds")

saveRDS(config_fit_stat_W, "rds/config_fit_stat_W.rds")
saveRDS(config_fit_stat_H, "rds/config_fit_stat_H.rds")
saveRDS(config_fit_stat, "rds/config_fit_stat.rds")

saveRDS(partial_fit_stat_W, "rds/partial_fit_stat_W.rds")
saveRDS(partial_fit_stat_H, "rds/partial_fit_stat_H.rds")
saveRDS(partial_fit_stat, "rds/partial_fit_stat.rds")
```

## Supporting analyses
 
### Latent correlation estimates 
```{r}
# Extract list of parameter estimates
est <- unnest_list(lavInspect(partial_fit, "est"))
# Compute correlation coefficients for each group (%o%: outer product)
lapply(est$psi, function(x) x / (sqrt(diag(x)) %o% sqrt(diag(x)))) 
```
### Census example: impact of measurement noninvariance
\mbox{} 
```{r message=FALSE}
# https://www2.census.gov/programs-surveys/demo/tables/hispanic-origin/1995/1994-cps/sumtab-1.txt

# First, compute approximate number of Hispanic individuals aged 10-19 using
# the percentages reported in Table 1 (rows 5-6, column `Hispanic-origin population`)
hisp10_19 <- 27521 * (9.1 + 8.5) / 100 * 1000

# Compare the proportion selected (PS) in the reference group (White) under partial
# invariance to the expected PS for the Hispanic group under partial invariance, after
# matching the underlying distributions

# PS for the Hispanic group assuming matched latent distributions
PS_H <- out_4g$summary[5, "E_R(Hispanic)"]

# PS for the reference group (benchmark under no measurement bias)
PS_ref <- out_4g$summary[5, "White"] 

round(PS_H - PS_ref, 2) # ~3% difference
ceiling(hisp10_19 * (PS_H - PS_ref)) # ~145,889 more individuals screened in
```

### Sensitivity analysis using alignment optimization

```{r message=FALSE}
library(sirt)

# Prepare the data and obtain the configural fit as before

# Extract loadings and intercepts
ests <- inspect(configural_fit, "est")
# Indices of different dimensions
ind_list <- list(1:6, 7:13, 14:17, 18:19)
# Group sizes
grp_size <- inspect(configural_fit, what = "nobs")
# First dimension
lam_fac1 <- lapply(ests, \(x) x$lambda[ind_list[[1]], 1])
lam_fac1 <- do.call(rbind, lam_fac1)
nu_fac1 <- lapply(ests, \(x) x$nu[ind_list[[1]], 1])
nu_fac1 <- do.call(rbind, nu_fac1)
align_fac1 <- invariance.alignment(
  lam_fac1,
  nu_fac1,
  wgt = matrix(sqrt(grp_size), nrow = length(grp_size), ncol = ncol(lam_fac1))
)
fac1_syntax <- paste0(
  "F1 ~~ c(", paste0(align_fac1$pars[, "psi0"]^2, collapse = ", "), ") * F1\n",
  "F1 ~ c(", paste0(align_fac1$pars[, "alpha0"], collapse = ", "), ") * 1\n"
)
# Second dimension
lam_fac2 <- lapply(ests, \(x) x$lambda[ind_list[[2]], 2])
lam_fac2 <- do.call(rbind, lam_fac2)
nu_fac2 <- lapply(ests, \(x) x$nu[ind_list[[2]], 1])
nu_fac2 <- do.call(rbind, nu_fac2)
align_fac2 <- invariance.alignment(
  lam_fac2,
  nu_fac2,
  wgt = matrix(sqrt(grp_size), nrow = length(grp_size), ncol = ncol(lam_fac2))
)
fac2_syntax <- paste0(
  "F2 ~~ c(", paste0(align_fac2$pars[, "psi0"]^2, collapse = ", "), ") * F2\n",
  "F2 ~ c(", paste0(align_fac2$pars[, "alpha0"], collapse = ", "), ") * 1\n"
)
# Third dimension
lam_fac3 <- lapply(ests, \(x) x$lambda[ind_list[[3]], 3])
lam_fac3 <- do.call(rbind, lam_fac3)
nu_fac3 <- lapply(ests, \(x) x$nu[ind_list[[3]], 1])
nu_fac3 <- do.call(rbind, nu_fac3)
align_fac3 <- invariance.alignment(
  lam_fac3,
  nu_fac3,
  wgt = matrix(sqrt(grp_size), nrow = length(grp_size), ncol = ncol(lam_fac3))
)
fac3_syntax <- paste0(
  "F3 ~~ c(", paste0(align_fac3$pars[, "psi0"]^2, collapse = ", "), ") * F3\n",
  "F3 ~ c(", paste0(align_fac3$pars[, "alpha0"], collapse = ", "), ") * 1\n"
)
# Fourth dimension
lam_fac4 <- lapply(ests, \(x) x$lambda[ind_list[[4]], 4])
lam_fac4 <- do.call(rbind, lam_fac4)
nu_fac4 <- lapply(ests, \(x) x$nu[ind_list[[4]], 1])
nu_fac4 <- do.call(rbind, nu_fac4)
align_fac4 <- invariance.alignment(
  lam_fac4,
  nu_fac4,
  wgt = matrix(sqrt(grp_size), nrow = length(grp_size), ncol = ncol(lam_fac4))
)
fac4_syntax <- paste0(
  "F4 ~~ c(", paste0(align_fac4$pars[, "psi0"]^2, collapse = ", "), ") * F4\n",
  "F4 ~ c(", paste0(align_fac4$pars[, "alpha0"], collapse = ", "), ") * 1\n"
)
# Rerun CFA with aligned latent means and variances
aligned_fit <- do.call(cfa, c(args_cfa, auto.fix.first = FALSE,
  model = paste0(m1, fac1_syntax, fac2_syntax, fac3_syntax, fac4_syntax)))
# summary(aligned_fit)
```

Examine $f_{\text{MACS}}$ effect size indices reflecting which items contribute the most bias.
```{r}
es_a <- pinsearch::pin_effsize(aligned_fit)
sort(round(es_a[, item_factor[item_factor %in% colnames(es_a)]], 3), 
     decreasing = TRUE)
```

\clearpage

# Tables

```{r include=FALSE}
build_model_fit_table(2, "rds/config_fit_stat_W.rds", "rds/partial_fit_stat_W.rds",
    suffix = "2g_W")
build_model_fit_table(2, "rds/config_fit_stat_H.rds", "rds/partial_fit_stat_H.rds",
    suffix = "2g_H")
```

```{r echo = FALSE, message = FALSE}
model_fit_kable(pth = c("rds/fit_tab2g_W.rds", "rds/fit_tab2g_H.rds"), 
                cap = " (two groups).", 
                cases = c("White vs. Non-White", "Non-Hispanic vs. Hispanic"),
                addft = "Groups were collapsed to achieve a binary grouping for illustrative purposes.")
```

\clearpage
```{r include=FALSE, eval = FALSE}
lapply(labs, function(x) 
  print(pin_out_W[[2]][pin_out_W[[2]]$lhs == x | pin_out_W[[2]]$rhs == x,]))
```

```{=latex}
\begin{table}
\centering
\caption{Non-invariant measurement parameters (two groups).}
\label{tab:noninvitemsh}
  \resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{threeparttable}
\begin{tabular}{l *{4}{>{\centering\arraybackslash}p{1cm}} | *{4}{>{\centering\arraybackslash}p{1cm}}}
\toprule
& \multicolumn{4}{c}{White vs. non-White} & \multicolumn{4}{c}{non-Hispanic vs. Hispanic}  \\\midrule
Item & $\lambda$ & $\nu$ & $\theta$ & $d_{\text{MACS}}$ & $\lambda$ & $\nu$ & $\theta$ & $d_{\text{MACS}}$
\\\midrule
\multicolumn{5}{l}{\emph{1. Somatic Complaints}} 
\\\midrule
 i1   (bothered)     & - & - & * & -    & - & * & * & .129    \\
 i2   (appetite)     & * & * & * & .092 & - & - & - & -       \\
 i5   (mind)         & - & - & - & -    & - & - & * & -         \\
 i7r  (tired)        & - & * & * & .097 & - & - & - & -         \\
 i13  (talk)         & - & * & * & .182 & - & - & * & -          \\
 i20r (getstarted)   & - & - & - & -        & - & - & - &  
 \\\hline
\multicolumn{5}{l}{\emph{2. Negative Affect}}
\\\midrule
 i3   (blues)        & - & - & * & -         & - & - & - & -    \\
 i6   (depressed)    & - & - & * & -         & - & - & - & -    \\
 i9   (failure)      & * & * & * & .246 & - & * & * & .148      \\
 i10  (fearful)      & * & * & * & .114 & - & * & * & .116      \\
 i14  (lonely)       & * & * & * & .114 & - & - & - & -         \\
 i18  (sad)          & - & - & * & -    & - & - & - & -         \\
 i21  (worthliving)  & * & * & * & .168 & - & * & * & .140     
 \\\hline
\multicolumn{5}{l}{\emph{3. (Lack of) Positive Affect}}
\\\midrule
 i4   (good)         & - & - & * & -    & - & * & * & .177      \\
 i8   (hopeful)      & - & - & * & -    & - & * & * & .126      \\
 i12  (happy)        & - & - & * & -    & - & - & * & -         \\
 i16  (enjoy)        & - & - & * & -    & - & - & - & -    
 \\\hline
\multicolumn{5}{l}{\emph{4. Interpersonal Relations}}
\\\midrule
 i15  (unfriendly)   & - & * & - & .087 & - & - & - & - \\
 i19  (dislike)      & - & - & * & -    & - & - & * & -    
 \\\bottomrule
\end{tabular}
\begin{tablenotes}
\item \textit{Note.} "-" indicates invariance. $\lambda$, $\nu$, and $\theta$ refer to factor loadings, intercepts, and uniqueness parameters. $d_{\text{MACS}}$ reflects the item-level effect size of measurement noninvariance across two groups. * indicates that the parameter was noninvariant.
\end{tablenotes}
\end{threeparttable}}
\end{table}
```

\clearpage
```{r include = FALSE}
build_parameter_table(2, "rds/pin_out_W.rds", c("White","non-White"),
                      suffix = "2g_W")
build_parameter_table(2, "rds/pin_out_H.rds", c("non-Hispanic", "Hispanic"),
                      suffix = "2g_H")
build_parameter_table(2, "rds/pin_out.rds", groups_wbha, suffix = "4g")
```

```{r echo = FALSE}
parameter_kable(
  pth = c("rds/parameter_tab2g_W.rds", "rds/parameter_tab2g_H.rds"),
  ng = 2, groups = list(c("NW", "W"), c("NH", "H")), 
  cases = c("White vs. Non-White", "Non-Hispanic vs. Hispanic"),
  addft = "NW, W, NH, and H refer to non-White, White, non-Hispanic, and Hispanic respectively. Groups were collapsed to achieve a binary grouping for illustrative purposes.")
```



\clearpage

```{r include = FALSE, eval = TRUE}
est_W <- unnest_list(lavInspect(pin_out_W[[1]], "est"))
est_H <- unnest_list(lavInspect(pin_out_H[[1]], "est"))
```

\clearpage
```{r echo = FALSE, message = FALSE}
build_psi_alpha_table(2, est_psi = est_W$psi, est_alpha = est_W$alpha, suffix = "2g_W")
psi_alpha_kable("rds/psi_alpha_tab2g_W.rds", groups = c("non-White", "White"))
```
\clearpage
```{r echo = FALSE, message = FALSE}
build_psi_alpha_table(2, est_psi = est_H$psi, est_alpha = est_H$alpha, suffix = "2g_H")
psi_alpha_kable("rds/psi_alpha_tab2g_H.rds", groups = c("non-Hispanic", "Hispanic"))
```
\clearpage
```{r echo = FALSE}
build_CAI_table(3, "rds/PartInv_out_4g.rds", ng = 4,suffix = "4g")
build_CAI_table(3, "rds/PartInv_out_2g_W.rds", ng = 2, suffix = "2g_W")
build_CAI_table(3, "rds/PartInv_out_2g_H.rds", ng = 2, suffix = "2g_H")
```
```{r echo = FALSE, message = FALSE}
CAI_kable(pth = c("rds/CAI_tab2g_W.rds", "rds/CAI_tab2g_H.rds"),
  groups = c("White","non-White", "non-Hispanic", "Hispanic"),
  cases = c("White vs Non-White", "Non-Hispanic vs. Hispanic"), ng = 2,
  cap = " (two groups)",
  addft = "Groups were collapsed to achieve a binary grouping for illustrative purposes.")
```

\clearpage
```{r echo=FALSE}
build_rEf_h_table("rds/PartInv_out_4g.rds", dig = 2, ng = 4, suffix = "4g")
build_rEf_h_table("rds/PartInv_out_2g_W.rds", dig = 2, ng = 2, suffix = "2g_W")
build_rEf_h_table("rds/PartInv_out_2g_H.rds", dig = 2, ng = 2, suffix = "2g_H")
```

```{r include = FALSE}
if (!file.exists("rds/href_tab4g_multi.rds")) {
  href_tab4g <- readRDS("rds/href_tab4g.rds")
  href_tab4g_multi <- cbind(href_tab4g, `$h_\\text{multi}$` = round(multi_h_4g, 2))
  saveRDS(href_tab4g_multi, "rds/href_tab4g_multi.rds")
}
```

```{r echo = FALSE}
rEf_h_kable(pth = c("rds/href_tab2g_W.rds", "rds/href_tab2g_H.rds"),
            groups = c(c("White", "non-White", "non-Hispanic", "non-Hispanic")),
            cases = c(c("White vs Non-White"), c("Non-Hispanic vs. Hispanic")),
            ng = 2,
  addft = " Groups were collapsed to achieve a binary grouping for illustrative purposes.")
```

\clearpage

# Figures 

```{=latex}
\begin{figure}[ht]
\caption{Joint bivariate distributions of observed and latent scores (White, non-White).}
\begin{minipage}{\textwidth}
  \begin{subfigure}{.5\textwidth}
    \includegraphics[width=1\textwidth]{figures/partial_W.png}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \includegraphics[width=1\textwidth]{figures/strict_W.png}
  \end{subfigure}
\end{minipage}
\parbox{\textwidth}{The figures illustrate the joint bivariate distributions of observed and latent scores for each group under partial (left) and strict (right) measurement invariance (two groups). Quadrants TP, FP, TN, and FN correspond to True Positive, False Positive, True Negative, and False Negative rates. Groups were collapsed to achieve a binary grouping for illustrative purposes.}
\end{figure}
```
\clearpage

```{=latex}
\begin{figure}[ht]
\caption{Joint bivariate distributions of observed and latent scores (non-Hispanic, Hispanic).}
\begin{minipage}{\textwidth}
  \begin{subfigure}{.5\textwidth}
    \includegraphics[width=1\textwidth]{figures/partial_H.png}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \includegraphics[width=1\textwidth]{figures/strict_H.png}
  \end{subfigure}
\end{minipage}
\parbox{\textwidth}{The figures illustrate the joint bivariate distributions of observed and latent scores for each group under partial (left) and strict (right) measurement invariance (two groups). Quadrants TP, FP, TN, and FN correspond to True Positive, False Positive, True Negative, and False Negative rates. Groups were collapsed to achieve a binary grouping for illustrative purposes.}
\end{figure}
```

\clearpage

```{=latex}
\begin{figure}[ht]
\caption{Proportions selected (PS) and success ratios (SR) at different cutoffs ($Z_c$) by group (White, non-White).}
\centering
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/PS_par_2g_W.png}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/PS_str_2g_W.png}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SR_par_2g_W.png}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SR_str_2g_W.png}
  \end{subfigure}
\parbox{\textwidth}{Figures on the left reflect partial MI and the figures on the right indicate strict MI. The vertical dotted line at 15.2 is equivalent to Radloff's (1977) suggested cutoff of 16 on 20 items. Groups were collapsed to achieve a binary grouping for illustrative purposes.}
\end{figure}
```
\clearpage
```{=latex}
\begin{figure}[ht]
\caption{Sensitivity (SE), Specificity (SP), and Adverse Impact ratios (AIR) at different cutoffs ($Z_c$) by group (White, non-White).}
\centering
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SE_par_2g_W.png}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SE_str_2g_W.png}
\end{subfigure}
\vspace{-0.5em}  
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SP_par_2g_W.png}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SP_str_2g_W.png}
\end{subfigure}
\vspace{-0.5em}
\begin{subfigure}{.5\textwidth}
\centering
  \includegraphics[width=\linewidth]{figures/AI_2g_W.png}
\end{subfigure}
\parbox{\textwidth}{In the first two rows, figures on the left reflect partial MI and the figures on the right indicate strict MI. The vertical dotted line at 15.2 is equivalent to Radloff's (1977) suggested cutoff of 16 on 20 items. Groups were collapsed to achieve a binary grouping for illustrative purposes.}
\end{figure}
```

\clearpage
```{=latex}
\begin{figure}[ht]
\caption{Proportions selected (PS) and success ratios (SR) at different cutoffs ($Z_c$) by group (non-Hispanic, Hispanic).}
\centering
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/PS_par_2g_H.png}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/PS_str_2g_H.png}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SR_par_2g_H.png}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SR_str_2g_H.png}
  \end{subfigure}
\parbox{\textwidth}{Figures on the left reflect partial MI and the figures on the right indicate strict MI. The vertical dotted line at 15.2 is equivalent to Radloff's (1977) suggested cutoff of 16 on 20 items. Groups were collapsed to achieve a binary grouping for illustrative purposes.}
\end{figure}
```

```{=latex}
\begin{figure}[ht]
\caption{Sensitivity (SE), Specificity (SP), and Adverse Impact ratios (AIR) at different cutoffs ($Z_c$) by group (non-Hispanic, Hispanic).}
\centering
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SE_par_2g_H.png}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SE_str_2g_H.png}
\end{subfigure}
\vspace{-0.5em}  
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SP_par_2g_H.png}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \includegraphics[width=\linewidth]{figures/SP_str_2g_H.png}
\end{subfigure}
\vspace{-0.5em}
\begin{subfigure}{.5\textwidth}
\centering
  \includegraphics[width=\linewidth]{figures/AI_2g_H.png}
\end{subfigure}
\parbox{\textwidth}{In the first two rows, figures on the left reflect partial MI and the figures on the right indicate strict MI. The vertical dotted line at 15.2 is equivalent to Radloff's (1977) suggested cutoff of 16 on 20 items. Groups were collapsed to achieve a binary grouping for illustrative purposes.}
\end{figure}
```

\clearpage

# Differential item weights

Here, we demonstrate how item weighting with factor scores can be implemented. In practice, sum scores are commonly used for instruments such as the CES-D, and this section is provided for illustrative purposes only. In the following, we used `lavPredict()` with the fitted `cfa` object to obtain Bartlett score coefficient matrices for each group, averaged these to derive a single set of coefficients, mapped items to their respective factors using the model syntax, and then extracted the corresponding item–factor coefficients to serve as item weights. The new cutoff adjusting for the revised weighting scheme was $z_c = 7.16$. \emph{pmix} was specified as before, and `PartInv()` was called with these configurations and the partial MI `cfa()` output from Step 1 to apply the multi-group MCAA framework.

```{r eval = FALSE, message=FALSE}
# Note: Before running the code below, run the relevant code chunks from the
# illustrative example to prepare the data, fit the configural and partial 
# models, define the mixing weights (mixp) etc.

# Latent factor weights: place weights proportional to item counts
lw <- c(6, 7, 4, 2) / sum(c(6, 7, 4, 2))

# Item weights: use Bartlett factor score coefficients
fsm <- attr(lavPredict(partial_fit, method = "bartlett", fsm = TRUE), "fsm") 
# Average matrices across groups 
fsm_avg <- Reduce("+", fsm) / length(fsm)
# Map items to factors using the model syntax string
map_i_F <- with(subset(lavaanify(m1), op == "=~"), setNames(lhs, rhs))
# Extract the coefficient for each item for the appropriate factor
iw <- fsm_avg[cbind(map_i_F, names(map_i_F))]
# adj_w: scaling constant = average of item weights (used to adjust cutoff values)
adj_w <- sum(iw) / 19 

# Call PartInv() using the cfa() output object
out_4g_w <- PartInv(
  cfa_fit = partial_fit, pmix = mixp, weights_latent = lw, weights_item = iw, 
  cut_z = 15.2 * adj_w, plot_contour = TRUE, show_mi_result = TRUE, custom_colors = cols)

# Plot CAI and Adverse Impact Ratio across a range of thresholds
plot_CAI_across_range(
  x = out_4g_w, custom_colors = cols, cutoffs_from = 10 * adj_w, 
  cutoffs_to = 30 * adj_w, add_vertical_threshold_at = 15.2 * adj_w)
```
